---
title: "Data Wrangling Steps"
author: "Anthony Stachowitz"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

### 1. Load datasets

This loads the individual datasets.   
```{r}
crime_between_1997_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_between_1997_2016.csv", header = TRUE, sep = ",")
crime_by_region <- read.csv("C:/Users/Anthony/cap_project/crime_by_region.csv", header = TRUE, sep = ",")
crime_city <- read.csv("C:/Users/Anthony/cap_project/crime_city.csv", header = TRUE, sep = ",")
crime_population_2015_vs_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_population_2015_vs_2016.csv", header = TRUE, sep = ",")
crime_population_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_population_2016.csv", header = TRUE, sep = ",")
```

### 2. Remove excess rows that contained only NA's in three of my data sets

remove rows with excess NA from all data sets crime_population_2016, crime_by_region, and crime_between_1997_2016. The NA's in this circumstance are just extra rows that were included during the transfer of data sets. They are not associated with missing or incorrect information and there is no need to convert them into data.
```{r}
crime_by_region <- crime_by_region[-c(199:208), ]
crime_population_2016 <- crime_population_2016[-c(15:24), ]
crime_between_1997_2016 <- crime_between_1997_2016[-c(21:42), ]
```

### 3. Remove specific values

Because my project looks at overall property crime rate and only looks at other types of crime when compared to the change in property crime over time, there are other crime values that need to be removed. For example, I am looking at the overall data concerning property crimes. Included in property crimes is data that is broken down by other types of crime. It is the other types of crime values that I have removed.
```{r}
crime_population_2016 <- crime_population_2016[,c(1:3,14,15,22,23)]
crime_by_region <- crime_by_region[,c(1:5,16,17)]
crime_between_1997_2016 <- crime_between_1997_2016[,c(1:4,15,16)]
crime_population_2015_vs_2016 <- crime_population_2015_vs_2016[,c(1:3,9,14,15)]
crime_city <- crime_city[,c(1:4,9)]
```

### 4. Change headers names to better reflect the data

There are some header names that are confusing and need to be more specific. For example, some of the information in one data set are the same as in a different data set but the header will not be consistent across the data sets. I have changed the headers so that they are all consistent.
```{r}
names(crime_population_2016) <- c("City_sizes", "Violent_offenses", "Rate_per_100,000", "Property_offenses", "Rate_per_100,000", "Total_number_areas", "Total_population")
names(crime_by_region) <- c("Region", "Year", "Population", "Violent_crime", "Rate_per_100,000", "Property_crime", "Rate_per_100,000")
names(crime_population_2015_vs_2016) <- c("Group_size", "Year", "Violent_crime", "Property_crime", "Number_of_groups", "2016_est_population")
names(crime_between_1997_2016) <- c("Year", "Population", "Violent_crime", "Rate_per_100,000", "Property_crime", "Rate_per_100,000")
names(crime_city) <- c("Metro_area", "Main_cities", "Population", "Violent_crime", "Property_crime")
```

### 5. Rearrange the data in the crime_city data set

The crime city data set is broken down by city metro area but the actual data is further broken down by city. The total of the city data for each metro area needed to me moved up to the rows with the metro area and then the city data needed to be removed. I accomplished that by breaking the data set into 2 separate data sets, cleaning them, and then recombining the data.
```{r}
CS_p1 <- filter(crime_city, !is.na(Metro_area))
CS_p2 <- filter(crime_city, Main_cities == "Total area actually reporting")
CS_p1 <- select(CS_p1, Metro_area, Population)
CS_p2 <- select(CS_p2, Population:Property_crime)
CS_main <- bind_cols(CS_p1, CS_p2)
```