---
title: "Data Wrangling Steps"
author: "Anthony Stachowitz"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

### 1. Load datasets

This loads the individual datasets.   
```{r}
crime_between_1997_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_between_1997_2016.csv", header = TRUE, sep = ",")
crime_by_region <- read.csv("C:/Users/Anthony/cap_project/crime_by_region.csv", header = TRUE, sep = ",")
crime_city <- read.csv("C:/Users/Anthony/cap_project/crime_city.csv", header = TRUE, sep = ",")
crime_population_2015_vs_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_population_2015_vs_2016.csv", header = TRUE, sep = ",")
crime_population_2016 <- read.csv("C:/Users/Anthony/cap_project/crime_population_2016.csv", header = TRUE, sep = ",")
```

### 2. Remove excess rows that contained only NA's in three of my data sets

remove rows with excess NA from all data sets crime_population_2016, crime_by_region, and crime_between_1997_2016.
```{r}
crime_by_region <- crime_by_region[-c(199:208), ]
crime_population_2016 <- crime_population_2016[-c(15:24), ]
crime_between_1997_2016 <- crime_between_1997_2016[-c(21:42), ]
```

### 3. Remove specific values

Because my project looks at overall property crime rate and only looks at other types of crime when compared to the change in property crime over time, there are other crime values that need to be removed.
```{r}
crime_population_2016 <- crime_population_2016[,c(1:3,14,15,22,23)]
crime_by_region <- crime_by_region[,c(1:5,16,17)]
crime_between_1997_2016 <- crime_between_1997_2016[,c(1:4,15,16)]
crime_population_2015_vs_2016 <- crime_population_2015_vs_2016[,c(1:3,9,14,15)]
crime_city <- crime_city[,c(1:4,9)]
```

### 4. Change headers names to better reflect the data

There are some header names that are confusing and need to be more specific.
```{r}
names(crime_population_2016) <- c("City_sizes", "Violent_offenses", "Rate_per_100,000", "Property_offenses", "Rate_per_100,000", "Total_number_areas", "Total_population")
names(crime_by_region) <- c("Region", "Year", "Population", "Violent_crime", "Rate_per_100,000", "Property_crime", "Rate_per_100,000")
names(crime_population_2015_vs_2016) <- c("Group_size", "Year", "Violent_crime", "Property_crime", "Number_of_groups", "2016_est_population")
names(crime_between_1997_2016) <- c("Year", "Population", "Violent_crime", "Rate_per_100,000", "Property_crime", "Rate_per_100,000")
names(crime_city) <- c("Metro_area", "Main_cities", "Population", "Violent_crime", "Property_crime")
```

